---
title: "Stat 133 Final Project"
author: "Sol Ah Yoo & Sharon Hong"
date: "December 9, 2015"
output: pdf_document
---
C:/Users/tjf00_000/OneDrive/Documents/finalproj
Import data 
```{r}
top_movie <- read.csv('/data/top_grossing_movie_yearly.csv')
```
Find the average total domestic box office by genre.
```{r}
avg <- aggregate(top_movie[, 4:5], list(top_movie$Genre), mean, na.rm = TRUE)
colnames(avg)[1] <- 'Genre'

top_movie$Genre <- as.factor(top_movie$Genre)

genre_freq <- count(top_movie, "Genre")
avg <- merge(avg, genre_freq)
avg <- avg[c(1, 4, 2, 3)]

colnames(avg)[3] <- "AvgProductionBudget"
colnames(avg)[4] <- "AvgDomesticBoxOffice"

g <- ggplot(avg, aes(x = Genre, y = freq))

g + geom_point(aes(size = AvgDomesticBoxOffice / 1000000,
                   color = AvgProductionBudget / 1000000)) +
  scale_size_continuous(range = c(10, 30))


```

How many different genres are there and what decade did they come out in?
What are the most popular genres?

```{r}
ggplot(top_movie, aes(x = Genre, fill = Decade)) + geom_bar() +
        labs(title = "Different Genres", y = "Count")
```

With the advancement of technology, directors are able to make realistic looking movies that require a lot of computer-generated imagery, such as adventure (ie. superhero).....?

Adventure movies tend to be more popular with the general public.


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
